---
title: "assignment"
output: html_document
date: "2026-03-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Using AnnotationHub, find and download the following annotations data


```{r}

library(AnnotationHub)
library(ensembldb)
library(GenomicRanges)
library(epiwraps)

```

1.1 The mouse (Mus Musculus) EnsDb object, version 102, genome build GRCm38

```{r}
ah <- AnnotationHub()

query(ah, c("EnsDb", "Mus musculus", "102", "GRCm38"))

mouse_ensdb_102 <- ah[["AH89211"]]

mouse_ensdb_102

```

1.2 The mouse genome sequence ( dna_sm ) in TwoBit/2bit format for GRCm38

```{r}
query(ah, c("Mus musculus", "GRCm38", "TwoBit"))

mouse_dnasm_2bit <- ah[["AH49775"]]

mouse_dnasm_2bit

```

1.3 The drosophila melanogaster genome sequence ( dna_sm ) in TwoBit/2bit format for BDGP6

```{r}

query(ah, c("Drosophila melanogaster", "BDGP6", "TwoBit"))

drosophila_dnasm_2bit <- ah[["AH106160"]]

drosophila_dnasm_2bit

```

2. Using the mouse EnsDb, find the following:

2.1 How many different ensembl gene IDs and gene symbols are there for protein-coding genes?

```{r}

mouse_genes <- genes(mouse_ensdb_102)

mouse_pcg <- mouse_genes[mouse_genes$gene_biotype=="protein_coding"]

length(mouse_pcg)
length(unique(mouse_pcg$gene_id))
length(unique(mouse_pcg$symbol))

```
2.2 Plot the distribution of the (spliced) length of protein-coding transcripts

```{r}

mouse_exons <- exonsBy(mouse_ensdb_102, filter = ~ tx_biotype == "protein_coding")
head(width(mouse_exons))

tx_lengths <- sum(width(mouse_exons))
head(tx_lengths)

```
```{r}

hist(tx_lengths,
     breaks=100,
     main = "Distribution of protein-coding transcript lengths",
     xlab = "Transcript length")

```

